Discovering causal factors with back-and-forth regression

### 2019-05-22 paper/paper.tex
better theorem, still in the making

\subsection{Asymptotic behaviour - linear case}
As discussed in the introduction, our problem can be formulated (up to a redefinition of the features of X) as $Y=f(EX+N)$, with F an unknown function.
%
If F is linear we can rewrite it as $Y = F(EX + N)$, with X an (n, dx) matrix of possible causes, E a (dx,dx) binary diagonal matrix selecting active features of X.
%
N a (dx,n) homoscedastic noise matrix, F an unknown (possibly full rank) (dy,dx) transformation corresponding to the measuring apparatus, and Y a (n,dy) matrix of measured effects.

@@

For the first step, replacing Y, and since N is centred, we have
\begin{equation}
\begin{aligned}
\arg \min_G \left \| X-GY \right \|^2 &= \arg \min_G \left \| X - GF(EX+N)\right\|^2 \\
&{}= \arg \min_G \left \| (I-GFE)X\right\| ^2  + \left \| GFN\right \| ^2\\
&{}= \arg \min_G \Tr (X'(I-GFE)'(I-GFE)X) + \Tr (N'F'G'GFN)\\
&{}= \arg \min_G \Tr ((I-GFE)'(I-GFE)XX') + \Tr (G'G(FN)(FN)')\\
\end{aligned}
\end{equation}

As noise N gets close to zero, and if F and X have full rank, then the solution tends to $(FE)^\dagger$, the pseudo inverse of FE.

@@

FC: something needed here about F and X'X not full rank or badly conditioned

>>>>

\subsection{Asymptotic behaviour - linear case}
As discussed in the introduction, our problem can be formulated (up to a redefinition of the features of X) as $Y=f(EX+N)$, with f an unknown function.
%
If F is linear we can rewrite it as $Y = F(EX + N)$, with X an (dx, n) matrix of possible causes, E a (dx,dx) binary diagonal matrix selecting active features of X.
%
N a (dx,n) homoscedastic noise matrix, F an unknown (possibly full rank) (dy,dx) transformation corresponding to the measuring apparatus, and Y a (dy, n) matrix of measured effects.

@@

For the first step, replacing Y, and since N is centred, we have
\begin{equation}
\begin{aligned}
\arg \min_G \left \| X-GY \right \|^2 &= \arg \min_G \left \| X - GF(EX+N)\right\|^2 \\
&{}= \arg \min_G \left \| X-GFEX\right\| ^2  + \left \| GFN\right \| ^2
\end{aligned}
\end{equation}

As N tends to zero, we are searching for the minimal norm solution of $\min_G \left \| X-GFEX\right\| ^2$, which is $X(FEX)^\dagger$ ($M^\dagger$ denoting the pseudo inverse of M).

Let k be the number of active features of E, suppose for notational convenience that they are the first k features of X, and let $FEX$ be of rank k (which means F and X are of full rank on the space spanned by the active features of E). Let K be the (nx,k) diagonal matrix built by deleting the rightmost nx-k columns of E. Then, $K=(\frac{I_{k}}{0})$. we may write $E=KK'$, and $FEX=(FK)(K'X)$, with $FK$  and $K'X$ of full rank k.

The rank factorisation property of the pseudo inverse (ref) says that, if $A=BC$, A (m,n), B(m,k) or rank k C(k,n) of rank k, then $A^\dagger=C^\dagger B^\dagger= C' (CC')^{-1}(B'B)^{-1}B'$.

Replacing in the expression of the minimal norm solution yields 
\begin{equation}
\begin{aligned}
X(FEX)^\dagger &=X((FK)(K'X))^\dagger \\
&=X(K'X)'((K'X)(K'X)')^{-1}((FK)(FK)')^{-1}(FK)' \\
&=(XX')K(K'XX'K)^{-1}(FKK'F')^{-1} K'F' 
\end{aligned}
\end{equation}

Let $XX' = \left(\begin{array}{c|c}\Sigma_{1} & \Sigma_{2} \\\hline \Sigma_{3} & \Sigma_{4}\end{array}\right)$, the top left block being (k,k), since $K=\left(\begin{array}{c}I \\\hline 0\end{array}\right)$, we have $(K'XX'K)=\Sigma_{1}$ and $(XX')K(K'XX'K)^{-1}=\left(\begin{array}{c}I_{k} \\\hline \Sigma_{3} \Sigma_{1}^{-1}\end{array}\right)=S$

Thus we have $X(FEX)^\dagger =S (FKK'F')^{-1} K'F'$

FC: this explains the negative diagonal elements, when X is negatively correlated. Need now to prove that this is S(FE)+

FC: onward to the reduction, I assume that we get a scaled value of S(FE)+, this needs a better proof

@@
